---
title: "HDNAextractions_analysis"
author: "Mary Grace Catapang"
date: '2022-06-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```
```{r}
library("tidyverse")
```
## Data Import
```{r col types}
col_dna <- cols (
  label = col_character(),
  catalog_number = col_character(),
  country = col_character(),
  year_collected = col_date(format = "%Y"),
  dna_extraction = col_character(),
  date_extracted = col_date("%d%b%y"),#tofix parsing error
  extracted_by = col_character(),
  overnight_incubation = col_character(),
  mass_g = col_double(),
  qubit = col_character(),
  concentration = col_double()
)
```
```{r initial data}
HDNA_data <- read_csv("./data/hDNA_extraction_NHRE_initialdata2.csv", col_types = col_dna)
```
## Plots Initial
This is just for data exploration and won't be going into the poster.
### Specimens
```{r country of origin}
ggplot(data = HDNA_data) + 
  geom_bar(mapping = aes(x = country, fill = contemporary)) +
  coord_flip()
```
```{r year of collection}
ggplot(data = HDNAinitdata) + 
  geom_bar(mapping = aes(x = year_collected, fill = country))
```

### Mass vs. Concentration 
```{r mass vs. concentration by country}
ggplot(HDNA_data, aes(x=mass_g, y=concentration, color=country)) +
  geom_point(size=2)
```
```{r mass vs. concentration by year}
ggplot(HDNA_data, aes(x=mass_g, y=concentration, color=year_collected)) +
  geom_point(size=2)
```
### Year vs. Concentration
```{r year vs. concentration scatterplot}
ggplot(HDNA_data, aes(x=year_collected, y=concentration, color=country)) + 
    geom_point(size=2)
```
## Map
A map showing the location of the specimens used for the analysis.
```{r load packages}
library(dplyr)
library(ggplot2)
library(viridis)
```
```{r data import}
trans_data <- read_csv("./data/hDNA_extraction_NHRE_data_transformed.csv")
```

```{r world map}
# world map, not clean
WorldMap <- map_data("world") %>%
               filter(region != "Antarctica") %>%
               fortify
```
```{r centroid for bubbles}
centroid <- read_csv("./data/centroid.csv")
```
```{r specimen data}
# set data types
cols_country <- cols(
  `Country` = col_character(),
  `longitude` = col_double(),
  `latitude` = col_double(),
  `ISO` = col_character(),
  `COUNTRYAFF` = col_character(),
  `AFF_ISO` = col_character(),
  `count_specimens` = col_integer()
)

# add geospatial data to specimen data
df1 <- trans_data %>% group_by(country) %>% summarise(count_specimens = n())
data_geo_country <- distinct(merge(centroid, df1, by = "country", all.y=TRUE, col_types = cols_country))
view(data_geo_country)
```
This map still requires fixing for aesthetics. Centroid data should also be modified to included places like Hawaii and Galapagos. It might be good to have a map centered in the Pacific?
```{r plotting map customized}
# Create breaks for the color scale
mybreaks <- c(1, 5, 12)

ggplot() +
  geom_polygon(data = WorldMap, aes(x=long, y = lat, group = group), fill="grey", alpha=0.9) +
  geom_point( data=data_geo_country, aes(x=long, y=lat, size=count_specimens, color=count_specimens, alpha = count_specimens))+
  scale_size_continuous(name="Number of Specimens", trans="log", range=c(3,10), breaks=mybreaks) +
    scale_alpha_continuous(name="Number of Specimens", trans="log", range=c(0.1, .9), breaks=mybreaks) +
    scale_color_viridis(option="magma", trans="log", breaks=mybreaks,name="Number of Specimens" ) +
  theme_void() + coord_map() + 
    guides( colour = guide_legend()) +
    ggtitle("Pocillopora DNA Extractions") +
    theme(
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#f5f5f2", color = NA), 
      panel.background = element_rect(fill = "#f5f5f2", color = NA), 
      legend.background = element_rect(fill = "#f5f5f2", color = NA),
      plot.title = element_text(size= 16, hjust=0.1, color = "#4e4d47", margin = margin(b = -0.1, t = 0.4, l = 2, unit = "cm")),
    )
```
## Correlation with specimen age 
There is no correlation between age of specimen and the DNA concentration collected.
```{r year vs. concentration scatterplot}
# linear trend + confidence interval
library(hrbrthemes)
ggplot(HDNA_data, aes(x=year_collected, y=concentration)) +
  geom_point() +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum()
```
```{r year vs. purity scatterplot}



```
## Statistical Analysis
I used t-test to identify whether there is a significant difference between the qubit concentration of the DNA extracted using different methods. A p-value of ≤ 0.05 is considered statistically signficant.

```{r data import}
BloodTissue <- filter(HDNA_data, contemporary == "no" && dna_extraction == "Qiagen Dneasy Blood & Tissue Kit" )
Autogen <- filter(HDNA_data, contemporary == "no" && dna_extraction == "Qiagen Dneasy Blood & Tissue Kit" )
PowerSoil <- read_csv("./data/pp_concentration.csv")
```
No significant statistical difference between Blood&Tissue and Autogen.
```{r t-test BloodTissue vs. Autogen}

t.test(BloodTissue, Autogen) 
```
There is significant statistical difference between Powersoil and Autogen.
```{r t-test PowerSoil vs. Autogen}

t.test(PowerSoil, Autogen) 
```
There is significant statistical difference between Blood&Tissue and Powersoil.
```{r t-test PowerSoil vs. BloodTissue}

t.test(PowerSoil, BloodTissue) 
```
## Boxplot
This plot still needs cleaning and editing to match poster aesthetics. 
```{r box plot concentration}
hDNA_extract <- filter(HDNA_data, contemporary == "no")
view(hDNA_extract)
ggplot(hDNA_extract, aes(x=as.factor(dna_extraction), y=concentration)) + 
    geom_boxplot(fill="slateblue", alpha=0.2) + 
    xlab("DNA extraction")
```
```{r box plot purity}



```

